### 2016-02-03 Manuscript/SW_early_life_telomeres_SOM.Rmd
Redraft edits

**Table S1** Model selection table for analysis of factors affecting early life telomere length in Seychelles warblers. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here. TQ = territory quality.

@@

#Run the model
lm1 <- lm(cenTL ~ Fledged*(cenHelper + cenNonHelper + Sex + cenTarsus  + cenTQ),
            data = juvall)

@@

#Sort out row/col names
nams <- colnames(t1)
nams[nams == 'cenTarsus'] <- 'Tarsus'
nams[nams == 'cenNonHelper'] <- 'Non helpers'
nams[nams == 'SexMales'] <- 'Sex'
nams[nams == 'cenTQ'] <- 'TQ'
nams[nams == 'cenHelper'] <- 'Helpers'
nams[nams == 'cenHelper:Fledged'] <- 'Ageclass x Helpers'
nams[nams == 'cenNonHelper:Fledged'] <- 'Ageclass x Non helpers'
nams[nams == 'cenTarsus:Fledged'] <- 'Ageclass x Tarsus'
nams[nams == 'cenTQ:Fledged'] <- 'Ageclass x TQ'
nams[nams == 'Fledged:SexMales'] <- 'Ageclass x Sex'
nams[nams == 'Fledged'] <- 'Ageclass'

@@

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:12)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

@@

>>>>

**Table S1** Model selection table for analysis of factors affecting early life telomere length in juvenile Seychelles warblers. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here. TQ = territory quality.

@@

#Run the model
lm1 <- lmer(LogTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|LayYear)+(1|PlateID)+(1|BirdID),
            data = juv,
            REML = FALSE)
lm1 <- standardize(lm1)

@@

#Sort out row/col names
nams <- colnames(t1)

nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

@@

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:14)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

@@

**Table S2** Model selection table for analysis of factors affecting early life telomere loss in Seychelles warbler chicks, using longitudinal data. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here. TQ = territory quality.

```{r Table S2}
temp <- subset(subset(Loss,Agemonths == 1),TimeDiff>0.3)

#Run the model
lm2 <- lmer(TROC ~ Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ + (1|LayYear),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

lm3 <- dredge(lm2)
lm3 <- subset(lm3,delta <=6)

#Model averaging
t1 <- as.matrix(subset(lm3, delta<=6))

#Sort out row/col names
nams <- colnames(t1)

nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:7)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

**Figure S1**
```{r Figure S1}
#####################
#4A
#####################
temp <- subset(subset(Loss,Agemonths == 1),TimeDiff>0.3)

#Run the model
lm2 <- lmer(TROC ~ Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ + (1|LayYear),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

lm3 <- dredge(lm2)
lm3 <- subset(lm3,delta <=6)
lmavg <- model.avg(lm3,fit=T)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table2 <- imptable(lmavg,myimp)
table2 <- signif(table2,2)

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.LogTL'] <- 'Initial TL'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

rownames(table2) <- nams
colnames(table2) <- c('Estimate','SE','2.5%','97.5%','RI')

table2 <- data.frame(table2,myx=row.names(table2))

FigS1 <- forestplot(table2)
FigS1

### 2016-02-03 Manuscript/SW_early_life_telomeres_tables_and_figures.Rmd
Redraft edits

**Figure 1** Telomere length (mean $\pm$ s.e.) in relation to age in Seychelles warblers, based on cross-sectional **(A)** and longitudinal **(B)** data. In **B**, telomere loss is a standardised and corrected rate (see Methods for details) based on the difference in telomere length in the sample taken from birds in their first year of life, and the first available sample as an adult. Values above zero indicate telomere shortening in early life.

**Figure 2** Temporal variation in early-life telomere dynamics in the Seychelles warbler based on cross-sectional **(A)** and longitudinal **(B)** data (see Fig. 1 and main text for details). Points and error bars represent mean and standard error of telomere length for all birds born in each main (black) or minor (grey) breeding season, in relation to island-wide food availability in that season. Lines and shaded areas represent fitted values and 95% confidence limits from a linear regression.

**Figure 3** Factors affecting early-life telomere length in Seychelles warblers at the individual level, within seasons. **A** Model averaged estimates and 95% confidence intervals for explanatory terms used in a linear model with telomere length as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; TQ = territory quality). **B,C** Raw telomere length Presence/absence of helpers (**B**) and tarsus length (**C**) in nestlings (black, solid lines) and fledglings (grey, dashed lines). Points and error bars in **B** are means $\pm$ s.e., and the line and shaded areas in **C** represent fitted values and 95% confidence limits from a linear regression.

**Figure 4** Temporal, cohort-level variation in lifespan in relation to early-life telomere dynamics in the Seychelles warbler, based on cross-sectional **(A)** and longitudinal **(B)** data (see Fig. 1 and main text for details). Points and error bars represent means and $\pm$ s.e., respectively, of all birds born in each main (black, solid lines) or minor (grey, dashed lines) breeding season. Lines and shaded areas represent fitted values and 95% confidence limits from a linear regression.

**Figure 5** Kaplan-Meier curves showing the relationship between survival and telomere length for individual Seychelles warblers measured as chicks (**A**), and fledglings  (**B**). Telomere length is binned into groups here for visualisation purposes only (long and short = greater than or less than median telomere length, respectively).

@@

####################
temp <- subset(juvseason,Season == 'Summer')
temp <- subset(temp,FieldPeriodID != 27)

Fig1 <- ggplot(temp,
       aes(x = CatchYear,
           y = TLKBmean,
           col = Season)) +
  ylab('Log telomere length (kb)') + 
  xlab('Lay Year') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  geom_line()+
  scale_color_manual(values = c('black',grey(0.5)))

################################

pdf('Figures/Figure 1.pdf',
    width = 6,
    height = 4,
    paper = 'a4')
Fig1
dev.off()

@@

####################
Fig1 <- ggplot(dd,
       aes(x = Fledged,
           y = LogTL,
           col = Season)) +
  ylab('Telomere length (kb)') + 
  xlab('Lay Year') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  geom_line()+
  scale_color_manual(values = c('black',grey(0.5)))

################################

pdf('Figures/Figure 1.pdf',
    width = 6,
    height = 4,
    paper = 'a4')
Fig1
dev.off()

@@

#####################

#Run the model
lm2 <- lmer(LogTL ~ Fledged + Tarsus + Helper + GroupSize + Sex +Tarsus + BodyMass  + cenTQ  + (1|LayYear) + (1|PlateID),
            data = juv,
            REML = FALSE)

#Model averaging

@@

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'cenTarsus'] <- 'Tarsus'
nams[nams == 'cenNonHelper'] <- 'Non helpers'
nams[nams == 'SexMales'] <- 'Sex'
nams[nams == 'cenTQ'] <- 'TQ'
nams[nams == 'cenHelper'] <- 'Helpers'
nams[nams == 'cenHelper:FledgedFledglings'] <- 'Ageclass x Helpers'
nams[nams == 'cenNonHelper:FledgedFledglings'] <- 'Ageclass x Non helpers'
nams[nams == 'cenTarsus:FledgedFledglings'] <- 'Ageclass x Tarsus'
nams[nams == 'cenTQ:FledgedFledglings'] <- 'Ageclass x TQ'
nams[nams == 'FledgedFledglings:SexMales'] <- 'Ageclass x Sex'
nams[nams == 'FledgedFledglings'] <- 'Ageclass'

@@

table2 <- data.frame(table2,myx=row.names(table2))

Fig3 <- forestplot(table2) +
  annotate('text',label = 'A',x=0.6,y=0.1)

pdf('Figures/Figure 2.pdf',
    width = 8,
    height = 8,
    paper = 'a4')
Fig3
dev.off()

@@

```{r Figure 3,fig.height=8,fig.width=4}

s1 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = chickall)
Fig3A <- ggsurv(s1,plot.cens=F,lty.est=c(1,2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('black',grey(0.5))) +
  guides(colour=F) +
  annotate('text',x=0.5,y=1.1,label='A')

###########################################
s2 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = FlSAall)
Fig3B <- ggsurv(s2,plot.cens=F,lty.est=c(1,2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('black',grey(0.5))) +
  guides(colour=F) +
  annotate('text',x=0.5,y=1.1,label='B')

pdf('Figures/Figure 3.pdf',
    width = 4,
    height = 8,
    paper = 'a4')
multiplot(Fig3A,Fig3B)
dev.off()

@@

>>>>

**Figure 1** Telomere length in relation to age in Seychelles warblers. **(A)** Mean  ($\pm$ s.e.) age for all birds. **(B)** Notched boxplot of telomere length at different age classes within the first year of life. Sample sizes for each group are provided in brackets.

**Figure 2** Temporal variation in early-life telomere length in the Seychelles warbler based on cross-sectional data. Points and error bars represent mean and standard error of telomere length for all birds born in each summer breeding season. Sample sizes for each year are provided in brackets.

**Figure 3** Factors affecting early-life telomere length in Seychelles warblers at the individual level, within seasons. Model averaged estimates and 95% confidence intervals for explanatory terms used in a linear mixed model with telomere length (cross-sectional data, **A**) and telomere loss (longitudinal data, **B**) as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; TQ = territory quality, TL = telomere length).

**Figure 4** Telomere length in relation to **A** season and **B** tarsus length in juvenile Seychelles warblers. In *A**, white and grey boxes represent summer and winter seasons, respectively, while in **B** black and grey points represent nestlings and fledglings, respectively. Sample sizes for each group are provided in brackets.

**Figure 5** Kaplan-Meier curves showing the relationship between survival and telomere length for juvenile Seychelles warblers. Telomere length is binned into groups here for visualisation purposes only (long and short = greater than or less than median telomere length, respectively).

@@

temp <- dd
temp$Agemonths[temp$Agemonths>10] <- '12+'
temp$Agemonths <- factor(temp$Agemonths,levels = c('1','6','10','12+'))

Fig1A <- ggplot(temp,aes(x = Agemonths,y = LogTL))+
  geom_jitter(col='grey') +
  ylab('Log telomere length (kb)') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none')+
  geom_boxplot(outlier.colour=NA)+
  annotate('text',x = 0.6,y = 4.8,label = 'A')

#
temp <- Loss
temp[,'Age (months)'] <- factor(temp$Agemonths)

Fig1B <- ggplot(temp,aes(x = LogTL,y = LogTL1,col = `Age (months)`,shape = `Age (months)`)) +
  ylab('Log adult telomere length (kb)') + 
  xlab('Log juvenile telomere length (kb)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none')+
  geom_point() +
  stat_smooth(method = 'lm')+
  xlim(2,4.5)+
  ylim(2,4.5)+
    scale_colour_manual(values = c('darkgrey','darkgreen','darkorange'))+
  annotate('text',x = 2,y = 4.5,label = 'B')

#
temp <- Loss
temp[,'Age (months)'] <- factor(temp$Agemonths)

Fig1C <- ggplot(temp,aes(x = TimeDiff,y = Loss,col=`Age (months)`,shape = `Age (months)`)) +
  ylab('Log telomere loss (kb)') + 
  xlab('Time interval (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none')+
  geom_point()+
  geom_abline(slope = 0,intercept = 0,lty = 2)+
  stat_smooth(method='lm')+
  scale_colour_manual(values = c('darkgrey','darkgreen','darkorange'))+
  annotate('text',x = 0,y = 2,label = 'C')

Fig1D <- ggplot(temp,aes(x = TimeDiff,y = TROC,col=`Age (months)`,shape = `Age (months)`)) +
  ylab('Log telomere loss rate (kb/year)') + 
  xlab('Time interval (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = c(0.9,0.8))+
  geom_point()+
  geom_abline(slope = 0,intercept = 0,lty = 2)+
  scale_colour_manual(values = c('darkgrey','darkgreen','darkorange'))+
  annotate('text',x = 0,y = 8,label = 'D')

pdf('Figures/Figure 1.pdf',
    width = 8,
    height = 8,
    paper = 'a4r')
multiplot(Fig1A,Fig1B,Fig1C,Fig1D,layout=matrix(c(1:4),nrow = 2,byrow=T))
dev.off()

@@

###############################
temp <- juv
temp$LayYear <- factor(temp$LayYear)

Fig2A <- ggplot(subset(temp,Agemonths == 1),
       aes(x = LayYear,
           y = LogTL)) +
  ylab('') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60,size = 14,vjust=0.5), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_jitter(col = 'grey')+
  ylim(c(2,4.5))+
  geom_boxplot(outlier.colour = NA)+
  scale_x_discrete(limits = factor(1998:2014),drop=F)+
  annotate('text',x = 1,y = 4.4,label = 'A')

Fig2B <- ggplot(subset(juv,Agemonths == 6),
       aes(x = factor(LayYear),
           y = LogTL)) +
  ylab('Log telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60,size = 14,vjust=0.5), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_jitter(col = 'grey')+
  ylim(c(2,4.5))+
  geom_boxplot(outlier.colour = NA)+
  scale_x_discrete(limits = factor(1998:2014),drop=F)+
  annotate('text',x = 1,y = 4.4,label = 'B')

Fig2C <- ggplot(subset(juv,Agemonths == 10),
       aes(x = factor(LayYear),
           y = LogTL)) +
  ylab('') + 
  xlab('Cohort') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60,size = 14,vjust=0.5), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_jitter(col = 'grey')+
  ylim(c(2,4.5))+
  geom_boxplot(outlier.colour = NA)+
  scale_x_discrete(limits = factor(1998:2014),drop=F)+
  annotate('text',x = 1,y = 4.4,label = 'C')

pdf('Figures/Figure 2.pdf',
    width = 6,
    height = 8,
    paper = 'a4r')
multiplot(Fig2A,Fig2B,Fig2C)
dev.off()

@@

temp <- juv

#####################
#3A
#####################

#Run the model
lm2 <- lmer(LogTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ)  + (1|LayYear) + (1|PlateID) + (1|TerritoryID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

@@

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

@@

table2 <- data.frame(table2,myx=row.names(table2))
table2$myx <- factor(table2$myx,levels = table2$myx[order(table2$Estimate,decreasing = T)])

Fig3 <- forestplot(table2)+
  ylim(-0.4,0.4)

pdf('Figures/Figure 3.pdf',
    width = 8,
    height = 5,
    paper = 'a4r')
Fig3
dev.off()

@@

```{r Figure 4}
temp <- juv

temp[,'Helpers'] <- ifelse(temp$Helper>0,'Present', 'Absent')
temp[,'Non helpers'] <- ifelse(temp$NonHelper>0,'Present', 'Absent')
temp[,'Age (months)'] <- factor(temp$Agemonths)

Fig4A <- ggplot(temp,
       aes(x = factor(Agemonths),
           y = LogTL,
           fill=`Non helpers`)) +
  ylab('Log telomere length (kb)') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = `Non helpers`),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 0.55,y = 4.5,label = 'A')

Fig4B <- ggplot(temp,
       aes(x = factor(Agemonths),
           y = LogTL,
           fill=`Helpers`)) +
  ylab('Log telomere length (kb)') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = `Helpers`),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 0.55,y = 4.5,label = 'B')

Fig4C <- ggplot(temp,
       aes(x = TQ,
           y = LogTL,
           colour=`Age (months)`,
           fill = `Age (months)`,
           shape = `Age (months)`)) +
  ylab('Log telomere length (kb)') + 
  xlab('Territory quality') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha = 0.2)+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))+
  annotate('text',x = 3,y = 4.5,label = 'C')

Fig4D <- ggplot(temp,
       aes(x = Tarsus,
           y = LogTL,
           colour=`Age (months)`,
           fill = `Age (months)`,
           shape = `Age (months)`)) +
  ylab('Log telomere length (kb)') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha = 0.2)+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))+
  annotate('text',x = 17,y = 4.5,label = 'D')

pdf('Figures/Figure 4.pdf',
    width = 7,
    height = 8,
    paper = 'a4r')
multiplot(Fig4A,Fig4B,Fig4C,Fig4D,layout=matrix(c(1:4),byrow = TRUE,nrow = 2))
dev.off()

@@

```{r Figure 5}
temp <- subset(juv,LayYear<2009)
mymed <- median(temp$cohortTL)
temp$coTLF <- ifelse(temp$cohortTL>mymed,'h','l')

s1 <- survfit(Surv(time = Lifespan, event = Died)~coTLF, data = temp)
Fig5A <- ggsurv(s1,plot.cens=F,lty.est=c(1:2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('grey','black')) +
  guides(colour=F)

mymed <- median(temp$cenTL)
temp$cenTLF <- ifelse(temp$cenTL>mymed,'h','l')
s1 <- survfit(Surv(time = Lifespan, event = Died)~cenTLF, data = temp)
Fig5B <- ggsurv(s1,plot.cens=F,lty.est=c(1,2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('black','grey')) +
  guides(colour=F)

pdf('Figures/Figure 5.pdf',
    width = 4,
    height = 8,
    paper = 'a4')
multiplot(Fig5A,Fig5B)
dev.off()
